---
title: "Introduction to the no.name R package"
author: "Claude Nadeau, Maikol Diasparra, Joel Barnes, Deirdre Hennessy, Rochelle Garner"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Introduction to the no.name R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
  img { width: 100%; height: auto; }
  p { margin-top: 1rem;line-height: 1.5rem; }
</style>

<img src="img/sig-blk-en.svg" style = "max-width: 300px; height: auto; margin-top: 48px; margin-bottom: 24px; border: 0;" alt="logo">

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This R package enables users to build box/compartment models via an Excel workbook. ODE (ordinary differential equation) or CTMC (continuous-time Markov chain) models can be built with either single (ODE, CTMC) or multiple age groups (ODE). This package also provides tools for users to perform sensitivity analsyes/parameter sweeps on their models and to visualize model results.

## Installation
### GitHub

The most recent version of `no.name` is available on GitHub: <a href="https://github.com/barnzilla/no.name" target="_blank">www.github.com/barnzilla/no.name</a>. The `devtools` package allows users to download and install `no.name` directly from GitHub.

```{r warning = FALSE, message = FALSE, eval = FALSE}
devtools::install_github("barnzilla/no.name", upgrade = "never")
library(no.name)
```

### Source package

Users can also download and install the source package using the `install.packages` function from the `utils` package, which comes pre-installed with R.

```{r warning = FALSE, message = FALSE, eval = FALSE}
utils::install.packages(
  pkgs = "https://github.com/barnzilla/r-package-sources/raw/master/no.name_12.04.tar.gz",
  repos = NULL, 
  type = "source"
)
library(no.name)
```

```{r warning = FALSE, message = FALSE, echo = FALSE}
library(no.name)
```
